---
title: "Combined Appendices - Paper 1 (Ideological Commitment Signaling via Non-State Armed Groups)"
subtitle: "Appendix 3: All Plots | Appendix 4: All Regression Tables"
format:
  html:
    toc: true
    number-sections: true
    theme: cosmo
    self-contained: true
  pdf:
    documentclass: report
    classoption: [a4paper, 12pt]
    toc: true
    number-sections: true
    fig-cap-location: top
    tbl-cap-location: top
---

```{r}
#| label: setup-combined-appendices
#| include: false

# =============================================================================
# combined-appendices.qmd - Single file containing ALL plots (App 3) + ALL tables (App 4) for Paper 1
# =============================================================================
library(here)
# Rule 7: every script/QMD uses here::i_am()
here::i_am("quarto/combined-appendices.qmd")
setwd(here::here())



# Create quarto subfolders and copy all files from results/
dir.create(here("quarto/plots"),  showWarnings = FALSE, recursive = TRUE)
dir.create(here("quarto/tables"), showWarnings = FALSE, recursive = TRUE)

# Copy plots
plot_files <- list.files(here("results/plots"), pattern = "\\.png$", full.names = TRUE)
for (f in plot_files) {
  file.copy(f, here("quarto/plots", basename(f)), overwrite = TRUE)
}

# Copy tables (stargazer .tex, modelsummary .md/.txt, coefficient .csv)
table_files <- list.files(here("results/tables"), pattern = "\\.(tex|txt|md|html|csv)$", full.names = TRUE)
for (f in table_files) {
  file.copy(f, here("quarto/tables", basename(f)), overwrite = TRUE)
}

library(knitr)
library(stargazer)
library(modelsummary)

message("Working directory forced to project root: ", getwd())
message("Copied ", length(plot_files), " plot files to quarto/plots/")
message("Copied ", length(table_files), " table files to quarto/tables/")
```

This single self-contained Quarto document produces **both** Appendix 3 (all plots) and Appendix 4 (all regression tables) for **Paper 1** in one output file.  
All files follow the project coding rules (Rules 3, 4, 5, 7, 9, 12, 14). Plots use friendly English labels and ordered X-axes; tables use `stargazer` (preferred) or `modelsummary`.

## Appendix 3: All Plots

```{r}
#| echo: false
#| message: false
#| fig-cap: "All plots generated for Paper 1 (Ideological Commitment Signaling) - ordered by filename"

img_folder <- "../results/plots"
img_files <- list.files(img_folder, pattern = "\\.png$", full.names = TRUE)
knitr::include_graphics(img_files)
```

## Appendix 4: All Regression Tables

### 4.1 Stargazer / LaTeX Tables (Journal-Ready)

```{r}
#| echo: false
#| results: asis

tex_files <- list.files(here("quarto/tables"), pattern = "\\.tex$", full.names = TRUE)
for (tf in sort(tex_files)) {
  cat("\n\\newpage\n")
  cat(readLines(tf), sep = "\n")
  cat("\n\n")
}
```

### 4.2 Modelsummary / Markdown Tables (HTML-Friendly)

```{r}
#| echo: false
#| results: asis

md_files <- list.files(here("quarto/tables"), pattern = "\\.md$", full.names = TRUE)
for (mf in sort(md_files)) {
  cat("\n\n")
  cat(readLines(mf), sep = "\n")
  cat("\n\n")
}
```

### 4.3 Coefficient CSV Extracts (Quick Reference)

```{r}
#| echo: false
#| results: asis

csv_files <- list.files(here("quarto/tables"), pattern = "\\.csv$", full.names = TRUE)
if (length(csv_files) > 0) {
  for (cf in sort(csv_files)) {
    cat("\n\n**", basename(cf), "**\n\n")
    tab <- read.csv(cf)
    knitr::kable(tab, format = "markdown", digits = 3, caption = basename(cf))
    cat("\n\n")
  }
}
```

```