# Appendix: Robustness Checks and Descriptive Statistics {#sec-appendix}

```{r}
#| label: setup
#| include: false

library(here)
here::i_am("docs/appendix.qmd")  # Ensure correct project root

source(here::here("R", "00_packages.R"))

```

```{r}


# Load cached dyad_ready for descriptive statistics
dyad_ready_path <- here::here("results", "dyad_ready.rds")
if (file.exists(dyad_ready_path)) {
  dyad_ready <- readRDS(dyad_ready_path)
} else {
  message("dyad_ready.rds not found. Run 02_data_prep.R first.")
}
# Load cached monadic_ready for leader-level descriptive stats
monadic_ready_path <- here::here("results", "monadic_ready.rds")
if (file.exists(monadic_ready_path)) {
  monadic_ready <- readRDS(monadic_ready_path)
} else {
  message("monadic_ready.rds not found. Run 02_data_prep.R first.")
}

```

```{r}

# Load all model objects
h1_models   <- readRDS(here::here("results", "h1_logit_models.rds"))
h2_models   <- readRDS(here::here("results", "h2_logit_models.rds"))
h3_models   <- readRDS(here::here("results", "h3_logit_models.rds"))
h4_models   <- readRDS(here::here("results", "h4_logit_models.rds"))
h5_logit_models <- readRDS(here::here("results", "h5_logit_models.rds"))
h6_logit_models <- readRDS(here::here("results", "h6_logit_models.rds"))
h8_init     <- readRDS(here::here("results", "h8_init_models.rds"))
h8_target   <- readRDS(here::here("results", "h8_target_models.rds"))
# Load H8 robustness models (optional - skip if file missing)
h8_robust <- tryCatch(
  readRDS(here::here("results", "h8_robust_models.rds")),
  error = function(e) {
    message("h8_robust_models.rds not found - skipping H8 robustness in appendix")
    NULL
  }
)
h9_survival <- readRDS(here::here("results", "h9_survival.rds"))

```


```{r}

# Universal coefficient extractor for any model type
# Avoids summary() which crashes on brglmFit and incomplete lm objects
extract_coefs <- function(model, model_name) {
  if (is.null(model)) return(NULL)
  tryCatch({
    if (inherits(model, "logistf")) {
      data.frame(
        model = model_name,
        term = names(coef(model)),
        estimate = coef(model),
        std.error = sqrt(diag(vcov(model))),
        p.value = model$prob,
        stringsAsFactors = FALSE, row.names = NULL
      )
    } else if (inherits(model, "coxph")) {
      s <- summary(model)
      data.frame(
        model = model_name,
        term = rownames(s$coefficients),
        estimate = s$coefficients[, "coef"],
        std.error = s$coefficients[, "se(coef)"],
        p.value = s$coefficients[, "Pr(>|z|)"],
        stringsAsFactors = FALSE, row.names = NULL
      )
    } else {
      # glm, brglmFit, lm - use coef()/vcov() directly
      beta <- coef(model)
      v <- tryCatch(vcov(model), error = function(e) NULL)
      if (is.null(v)) {
        # If vcov fails, try to get SEs from model$coefficients matrix
        se <- rep(NA_real_, length(beta))
      } else {
        se <- sqrt(diag(v))
      }
      z <- beta / se
      pval <- 2 * pnorm(abs(z), lower.tail = FALSE)
      data.frame(
        model = model_name,
        term = names(beta),
        estimate = beta,
        std.error = se,
        p.value = pval,
        stringsAsFactors = FALSE, row.names = NULL
      )
    }
  }, error = function(e) {
    message("Could not extract from ", model_name, ": ", e$message)
    NULL
  })
}

```


```{r}

# Helper to build a kable comparison table from a named list of models
build_model_table <- function(model_list, coef_labels = NULL) {
  all_coefs <- do.call(rbind, Filter(Negate(is.null),
    Map(extract_coefs, model_list, names(model_list))
  ))
  if (is.null(all_coefs) || nrow(all_coefs) == 0) {
    message("No model coefficients could be extracted.")
    return(invisible(NULL))
  }
  if (!is.null(coef_labels)) {
    all_coefs <- all_coefs[all_coefs$term %in% names(coef_labels), ]
    all_coefs$label <- coef_labels[all_coefs$term]
  } else {
    all_coefs$label <- all_coefs$term
  }
  all_coefs$stars <- ifelse(all_coefs$p.value < 0.001, "***",
    ifelse(all_coefs$p.value < 0.01, "**",
      ifelse(all_coefs$p.value < 0.05, "*", "")))
  all_coefs$cell <- paste0(sprintf("%.3f", all_coefs$estimate),
    all_coefs$stars, " (", sprintf("%.3f", all_coefs$std.error), ")")
  wide <- reshape(all_coefs[, c("label", "model", "cell")],
    idvar = "label", timevar = "model", direction = "wide")
  names(wide) <- gsub("cell\\.", "", names(wide))
  knitr::kable(wide, row.names = FALSE)
}
```
This appendix presents supplementary material, including descriptive statistics for all key variables, robustness checks with alternative model specifications, and additional interaction plots.

## A.1 Descriptive Statistics

```{r}
#| label: tbl-descriptive
#| tbl-cap: "Descriptive Statistics for Key Variables (Full Dyadic Sample)"
key_vars <- dyad_ready %>%
  dplyr::select(
    mid_initiated,
    targets_democracy,
    sidea_revisionist_domestic,
    legit_ratio,
    v2exl_legitideol_a,
    v2exl_legitperf_a,
    v2exl_legitlead_a,
    sidea_winning_coalition_size,
    cinc_a,
    sidea_religious_support,
    sidea_party_elite_support,
    sidea_military_support,
    sidea_dynamic_leader,
    cold_war
  )
datasummary_skim(
  key_vars,
  type = "numeric",
  title = "Descriptive Statistics: Key Dyadic Variables",
  notes = "Autocratic Side A dyad-years, 1946\u20132010."
)
```

```{r}
#| label: tbl-descriptive-mono
#| tbl-cap: "Descriptive Statistics: Leader-Level Variables (Monadic Sample)"
if (exists("monadic_ready")) {
  mono_vars <- monadic_ready %>%
    dplyr::select(
      tenure_years,
      survived_to_end,
      avg_ideological_legit,
      legit_ratio,
      conflicts_vs_ideo_targets,
      conflicts_vs_democracy,
      avg_autocracy_level,
      log_avg_gdp,
      irregular_entry
    )
  datasummary_skim(
    mono_vars,
    type = "numeric",
    title = "Descriptive Statistics: Leader-Level (Monadic) Variables",
    notes = "Autocratic leader-years, 1946\u20132010."
  )
}
```

## A.2 Correlation Matrix

```{r}
#| label: fig-corr
#| fig-cap: "Correlation matrix of key explanatory variables"
library(ggcorrplot)
corr_data <- dyad_ready %>%
  dplyr::select(
    sidea_revisionist_domestic,
    legit_ratio,
    v2exl_legitperf_a,
    v2exl_legitlead_a,
    sidea_winning_coalition_size,
    sidea_religious_support,
    sidea_military_support,
    sidea_dynamic_leader
  ) %>%
  drop_na()

corr_matrix <- cor(corr_data)

ggcorrplot(
  corr_matrix,
  hc.order = TRUE,
  type = "lower",
  lab = TRUE,
  lab_size = 3,
  ggtheme = ggplot2::theme_minimal()
)
```

## A.3 Robustness: H1–H4 with Rare Events Logit (Firth)

We re-estimate H1–H4 using Firth's penalized likelihood correction for rare events. This addresses potential bias introduced by the low proportion of MID initiations in the full dyadic sample.

```{r}
#| label: tbl-rare-events
#| tbl-cap: "H1\u2013H4 Rare Events Robustness: Firth Penalized Logistic Regression"
library(logistf)

firth_h1 <- tryCatch(
  logistf(
    mid_initiated ~ sidea_revisionist_domestic +
      log_cinc_a + log_cinc_b +
      sidea_winning_coalition_size + t + cold_war,
    data = dyad_ready
  ),
  error = function(e) { message("Firth H1 failed: ", e$message); NULL }
)

firth_h3 <- tryCatch(
  logistf(
    mid_initiated ~ legit_ratio +
      log_cinc_a + log_cinc_b +
      sidea_winning_coalition_size + t + cold_war,
    data = dyad_ready
  ),
  error = function(e) { message("Firth H3 failed: ", e$message); NULL }
)

build_model_table(
  list(
    "H1 (Firth)" = firth_h1,
    "H1 (Logit)" = h1_models$h1_full,
    "H3 (Firth)" = firth_h3,
    "H3 (Logit)" = h3_models$h3_full
  ),
  coef_labels = c(
    sidea_revisionist_domestic = "Revisionist Ideology",
    legit_ratio = "Legitimation Ratio",
    log_cinc_a = "Log CINC (A)",
    log_cinc_b = "Log CINC (B)",
    sidea_winning_coalition_size = "Winning Coalition",
    cold_war = "Cold War",
    t = "Peace Years"
  )
)
```

## A.4 Robustness: H8 Moderation with Alternative Interactions

```{r}
#| label: tbl-h8-robust
#| tbl-cap: "H8 Robustness: Alternative Interaction Models (LPM)"
if (!is.null(h8_robust)) {
  build_model_table(
    list(
      "W. Coalition" = h8_robust$h8_wcoal_int,
      "Mil. Support" = h8_robust$h8_milsupp_int,
      "Cold War" = h8_robust$h8_coldwar_int
    )
  )
} else {
  cat("H8 robustness models not available (file missing or skipped).")
}
```

## A.5 H9 Robustness: Alternative Survival Specifications

```{r}
#| label: tbl-h9-robust
#| tbl-cap: "H9 Robustness: Full Cox PH Model and Interaction Specifications"
build_model_table(
  list(
    "Full Model" = h9_survival$cox_h9_full,
    "Int. Ideology" = h9_survival$cox_h9_int_ideology,
    "Int. Ratio" = h9_survival$cox_h9_int_ratio
  ),
  coef_labels = c(
    conflicts_vs_ideo_targets = "Conflict vs. Ideo. Target",
    conflicts_vs_democracy = "Conflict vs. Democracy",
    avg_ideological_legit = "Avg. Ideo. Legitimation",
    legit_ratio = "Legitimation Ratio",
    avg_autocracy_level = "Autocracy Level",
    log_avg_gdp = "Log GDP",
    log_avg_pop = "Log Pop",
    irregular_entry = "Irregular Entry"
  )
)
```

## A.6 Variable Coding Details

The table below details the exact coding decisions for binary and continuous variables used across all model tiers.

| Variable | Coding | Source |
|---|---|---|
| `mid_initiated` | 1 if `hihosta >= 2` (hostility level at or above threat of force) | GRAVE-D |
| `targets_democracy` | 1 if `v2x_libdem_b >= 0.5` (liberal democracy score of target) | GRAVE-D + V-Dem |
| `sidea_revisionist_domestic` | Composite GRAVE-D revisionist ideology index (0–1) | GRAVE-D |
| `legit_ratio` | `v2exl_legitideol_a / legit_total`; `legit_total = sum(ideol, lead, perf)` | V-Dem |
| `sidea_winning_coalition_size` | V-Dem/BdM coalition size proxy (0–1) | V-Dem |
| `sidea_dynamic_leader` | Binary indicator: dynamic/charismatic leader (GRAVE-D coding) | GRAVE-D |
| `log_cinc_a` / `log_cinc_b` | Natural log of CINC score for Side A / Side B | COW |
| `cold_war` | 1 if year <= 1991 | Derived |
| `t`, `t2`, `t3` | Peace years since last MID; polynomial splines for temporal dependence | Derived |

: Variable Coding Details {#tbl-variable-coding}

## A.7 Robustness: Full Models as Linear Probability Models (LPM) with Clustered SEs

To address potential functional form concerns and incidental parameters bias, we re-estimate the full models as linear probability models with clustered standard errors (on dyad).

```{r}
#| label: tbl-lpm-appendix
#| tbl-cap: "Appendix: Linear Probability Models – Full Specifications (Clustered SEs)"
lpm_models <- readRDS(here::here("results", "lpm_appendix_models.rds"))

build_model_table(
  list(
    "H1 Full (LPM)" = lpm_models$h1_full,
    "H2 Full (LPM)" = lpm_models$h2_full,
    "H3 Full (LPM)" = lpm_models$h3_full,
    "H4 Full (LPM)" = lpm_models$h4_full,
    "H5 Full (LPM)" = lpm_models$h5_full,
    "H6 Full (LPM)" = lpm_models$h6_full
  )
)

```