# Appendix: Robustness Checks and Descriptive Statistics {#sec-appendix}

```{r}
#| label: setup
#| include: false
source(here::here("R", "00_packages.R"))
source(here::here("R", "02_data_prep.R"))

# Load all model objects
h1_models   <- readRDS(here::here("results", "h1_logit_models.rds"))
h2_models   <- readRDS(here::here("results", "h2_logit_models.rds"))
h3_models   <- readRDS(here::here("results", "h3_logit_models.rds"))
h4_models   <- readRDS(here::here("results", "h4_logit_models.rds"))
h4_unfiltered <- readRDS(here::here("results", "h4_unfiltered_logit_models.rds"))
h5_logit_models <- readRDS(here::here("results", "h5_logit_models.rds"))
h6_logit_models <- readRDS(here::here("results", "h6_logit_models.rds"))
h8_init     <- readRDS(here::here("results", "h8_init_models.rds"))
h8_target   <- readRDS(here::here("results", "h8_target_models.rds"))
h8_robust   <- readRDS(here::here("results", "h8_robust_models.rds"))
h9_survival <- readRDS(here::here("results", "h9_survival.rds"))
```

This appendix presents supplementary material, including descriptive statistics for all key variables, robustness checks with alternative model specifications, and additional interaction plots.

## A.1 Descriptive Statistics

```{r}
#| label: tbl-descriptive
#| tbl-cap: "Descriptive Statistics for Key Variables (Full Dyadic Sample)"
key_vars <- dyad_ready %>%
  select(
    mid_initiated,
    targets_democracy,
    sidea_revisionist_domestic,
    legit_ratio,
    v2exl_legitideol_a,
    v2exl_legitperf_a,
    v2exl_legitlead_a,
    sidea_winning_coalition_size,
    sidea_national_military_capabilities,
    sidea_religious_support,
    sidea_party_elite_support,
    sidea_military_support,
    sidea_dynamic_leader,
    cold_war
  )

datasummary_skim(
  key_vars,
  type = "numeric",
  title = "Descriptive Statistics: Key Dyadic Variables",
  notes = "Autocratic Side A dyad-years, 1946\u20132010."
)
```

```{r}
#| label: tbl-descriptive-mono
#| tbl-cap: "Descriptive Statistics: Leader-Level Variables (Monadic Sample)"
if (exists("monadic_ready")) {
  mono_vars <- monadic_ready %>%
    select(
      tenure_years,
      survived_to_end,
      avg_ideological_legit,
      legit_ratio,
      conflicts_vs_ideo_targets,
      conflicts_vs_democracy,
      avg_autocracy_level,
      log_avg_gdp,
      irregular_entry
    )

  datasummary_skim(
    mono_vars,
    type = "numeric",
    title = "Descriptive Statistics: Leader-Level (Monadic) Variables",
    notes = "Autocratic leader-years, 1946\u20132010."
  )
}
```

## A.2 Correlation Matrix

```{r}
#| label: fig-corr
#| fig-cap: "Correlation matrix of key explanatory variables"
library(ggcorrplot)

corr_data <- dyad_ready %>%
  select(
    sidea_revisionist_domestic, legit_ratio,
    v2exl_legitperf_a, v2exl_legitlead_a,
    sidea_winning_coalition_size,
    sidea_religious_support, sidea_military_support,
    sidea_dynamic_leader
  ) %>%
  drop_na()

corr_matrix <- cor(corr_data)

ggcorrplot(
  corr_matrix,
  hc.order  = TRUE,
  type      = "lower",
  lab       = TRUE,
  lab_size  = 3,
  ggtheme   = ggplot2::theme_minimal()
)
```

## A.3 Robustness: H1–H4 with Rare Events Logit (Firth)

We re-estimate H1–H4 using Firth's penalized likelihood correction for rare events. This addresses potential bias introduced by the low proportion of MID initiations in the full dyadic sample.

```{r}
#| label: tbl-rare-events
#| tbl-cap: "H1\u2013H4 Rare Events Robustness: Firth Penalized Logistic Regression"
library(logistf)

firth_h1 <- logistf(
  mid_initiated ~ sidea_revisionist_domestic +
    log_cinc_a + log_cinc_b +
    sidea_winning_coalition_size + t + cold_war,
  data = dyad_ready
)

firth_h3 <- logistf(
  mid_initiated ~ legit_ratio +
    log_cinc_a + log_cinc_b +
    sidea_winning_coalition_size + t + cold_war,
  data = dyad_ready
)

# Results compared to main logit estimates
modelsummary(
  list(
    "H1 (Firth)"  = firth_h1,
    "H1 (Logit)"  = h1_models$h1_full,
    "H3 (Firth)"  = firth_h3,
    "H3 (Logit)"  = h3_models$h3_full
  ),
  coef_map = c(
    sidea_revisionist_domestic = "Revisionist Ideology",
    legit_ratio                = "Legitimation Ratio",
    log_cinc_a                 = "Log CINC (A)",
    sidea_winning_coalition_size = "Winning Coalition",
    cold_war                   = "Cold War"
  ),
  stars = c("*" = 0.05, "**" = 0.01, "***" = 0.001),
  gof_omit = "AIC|BIC|Log"
)
```

## A.4 Robustness: H8 Moderation with Alternative Interactions

```{r}
#| label: tbl-h8-robust
#| tbl-cap: "H8 Robustness: Alternative Interaction Models (LPM)"
#| results: asis
stargazer(
  h8_robust$h8_wcoal_int,
  h8_robust$h8_milsupp_int,
  h8_robust$h8_coldwar_int,
  type = "html",
  title = "H8 Robustness: Alternative Interaction Models",
  dep.var.labels = "Targets Democracy (LPM)",
  column.labels = c("W. Coalition", "Mil. Support", "Cold War"),
  star.cutoffs = c(0.05, 0.01, 0.001),
  omit.stat = c("aic", "ll"),
  notes = c("Sample: Autocratic MID initiators only.",
            "*** p<0.001, ** p<0.01, * p<0.05")
)
```

## A.5 H9 Robustness: Alternative Survival Specifications

```{r}
#| label: tbl-h9-robust
#| tbl-cap: "H9 Robustness: Full Cox PH Model and Interaction Specifications"
#| results: asis
stargazer(
  h9_survival$cox_h9_full,
  h9_survival$cox_h9_int_ideology,
  h9_survival$cox_h9_int_ratio,
  type = "html",
  title = "H9 Robustness: Additional Cox PH Specifications",
  dep.var.labels = "Hazard of Leader Removal",
  column.labels = c("Full Model", "Int. Ideology", "Int. Ratio"),
  covariate.labels = c(
    "Conflict vs. Ideo. Target",
    "Conflict vs. Democracy",
    "Avg. Ideo. Legitimation",
    "Legitimation Ratio",
    "Conflict × Ideo. Legit.",
    "Conflict × Legit. Ratio",
    "Autocracy Level", "Log GDP", "Log Pop", "Irregular Entry"
  ),
  star.cutoffs = c(0.05, 0.01, 0.001),
  omit.stat = c("aic", "ll")
)
```

## A.6 Variable Coding Details

The table below details the exact coding decisions for binary and continuous variables used across all model tiers.

| Variable | Coding | Source |
|---|---|---|
| `mid_initiated` | 1 if `hihosta >= 2` (hostility level at or above threat of force) | GRAVE-D |
| `targets_democracy` | 1 if `v2x_libdem_b >= 0.5` (liberal democracy score of target) | GRAVE-D + V-Dem |
| `sidea_revisionist_domestic` | Composite GRAVE-D revisionist ideology index (0–1) | GRAVE-D |
| `legit_ratio` | `v2exl_legitideol_a / legit_total`; `legit_total = sum(ideol, lead, perf)` | V-Dem |
| `sidea_winning_coalition_size` | V-Dem/BdM coalition size proxy (0–1) | V-Dem |
| `sidea_dynamic_leader` | Binary indicator: dynamic/charismatic leader (GRAVE-D coding) | GRAVE-D |
| `log_cinc_a` / `log_cinc_b` | Natural log of CINC score for Side A / Side B | COW |
| `cold_war` | 1 if year <= 1991 | Derived |
| `t`, `t2`, `t3` | Peace years since last MID; polynomial splines for temporal dependence | Derived |

: Variable Coding Details {#tbl-variable-coding}
