# Data and Methods {#sec-data-methods}

```{r}
#| label: data_methods_setup
#| include: false
here::i_am("docs/data_methods.qmd")  # Ensure correct project root

source(here::here("R", "00_packages.R"))

# Load cached dyad_ready (saved by 02_data_prep.R) instead of full pipeline
dyad_ready_path <- here::here("results", "dyad_ready.rds")
if (file.exists(dyad_ready_path)) {
  dyad_ready <- readRDS(dyad_ready_path)
} else {
  message("dyad_ready.rds not found. Run 02_data_prep.R first to generate it.")
}
```

## Data Sources

The primary dataset is **GRAVE_D_Master_with_Leaders.csv**, produced by the GRAVE-D data pipeline (Hanna 2026). GRAVE-D is not itself a primary data collection effort; rather, it is an integration, harmonization, and extension layer built on top of more than a dozen established political science and economics datasets. Its principal contributions are:

- **Structural organization**: converting monadic country-year source data into a directed dyad-year panel using the Formal Bilateral Influence Capacity (FBIC) index [@MoyerEtAl_FBIC] as the dyadic spine, enabling analysis of bilateral relationships;
- **Derived variables**: constructing composite indices (e.g., *revisionist potential*, *ideological extremity*, *religious distance*) and interaction terms not present in any single source;
- **Missingness management**: applying a multi-stage imputation protocol (linear interpolation, regional-year median fill, and derived variable recomputation) documented in the GRAVE-D codebook;
- **Original coding**: ALBA (Bolivarian Alliance) membership indicators.

The underlying source datasets, with their required citations, are:

1. **Formal Bilateral Influence Capacity (FBIC) Index** [@MoyerEtAl_FBIC]: Provides the directed dyad-year spine and all bandwidth variables (total, economic, political, security).

2. **Varieties of Democracy (V-Dem)** v15 [@VDem_Dataset]: Supplies democracy and regime indices, legitimation strategy variables (`v2exl_legitideol`, `v2exl_legitlead`, `v2exl_legitperf`, `v2exl_legitratio`), civil society, corruption, rule of law, power distribution, regime support group, and head-of-state/government variables for both sides of each dyad.

3. **V-Dem Episodes of Regime Transformation (ERT)** [@EdgellEtAl2020_ERT]: Regime transition episodes data.

4. **Correlates of War --- Militarized Interstate Disputes (MIDs)** v5 [@COW_MIDs_v5]: Provides hostility-level data from which `mid_initiated` and conflict outcomes are derived.

5. **Correlates of War --- National Material Capabilities (NMC)** v6 [@COW_NMC_v6]: Provides Composite Index of National Capability (CINC) scores used as controls for material power.

6. **Alliance Treaty Obligations and Provisions (ATOP)** v5.1 [@LeedsEtAl2002_ATOP; @ATOP_Dataset]: Alliance type indicators (defense, offense, neutrality, non-aggression, consultation).

7. **Correlates of War --- World Religion Project** [@COW_WorldReligion]: Religious demographic shares used to compute religious distance variables.

8. **Archigos** [@GoemansEtAl2009_Archigos; @Archigos_Dataset]: Leader tenure, entry/exit modes, and post-tenure fate variables.

9. **Colgan Revolutionary Leaders Dataset** [@Colgan2013_RevolutionaryLeaders; @Colgan_RevolutionaryLeaders_Dataset]: Revolutionary leader classification and revolutionary change category variables.

10. **Global Leader Ideology Dataset** [@Herre2022_GlobalLeaderIdeology_Dataset]: Leader ideology scores, legal tradition, and polity-level controls.

11. **Polity5**: Polity democracy/autocracy scores and regime durability.

12. **First Use of Violent Force Dataset** [@CaprioliTrumbore2003_FirstUse]: First-use-of-force coding for MIDs.

13. **Hanson and Sigman State Capacity Index** [@HansonSigman2021_StateCapacity]: State fiscal and administrative capacity measures.

14. **Fraser Institute Economic Freedom Dataset** [@GwartneyEtAl2022_EFW_Dataset]: Economic freedom indices.

15. **Maddison Project Database** [@BoltVanZanden2014_MaddisonProject; @MaddisonProject_Database]: Historical GDP per capita.

16. **Ross Oil and Gas Dataset** [@Ross_OilGas_Dataset]: Fuel and oil income per capita.

17. **Standardized World Income Inequality Database (SWIID)** [@Solt2020_SWIID; @SWIID_Dataset]: Income inequality (Gini) estimates.

18. **RED --- Relational Export Dataset** [@RED_Dataset]: Trade relationship data.

19. **ALBA Membership** (Hanna 2026): Original coding of Bolivarian Alliance membership.

The sample is restricted to dyad-years in which Side A is classified as autocratic ($v2x\_libdem_a < 0.5$) and the year is 1946 or later (post-World War II international order).

## Variables

### Dependent Variables

**`mid_initiated`** (Tier 1–2 analyses): Binary indicator equal to 1 when Side A initiates a Militarized Interstate Dispute, defined as `hihosta >= 2` (threat or use of force). Constructed in `02_data_prep.R`.

**`targets_democracy`** (Tier 1–2 analyses): Binary indicator equal to 1 when the dyad's Side B is classified as democratic ($v2x\_libdem_b \geq 0.5$), conditional on MID initiation. For H2 and H4 models this variable is the dependent variable on the subset of initiator-years.

**Leader tenure / exit hazard** (Tier 3 survival analysis): Measured in leader-years, with exit coded as an event indicator for Cox Proportional Hazard models.

### Primary Independent Variables

**`sidea_revisionist_domestic`** (H1, H2): GRAVE-D composite measure of leadership revisionist ideology for Side A's governing leadership. Higher values indicate stronger revisionist ideological orientation.

**Support group variables** (H3, H4): GRAVE-D indicators for the regime's reliance on specific ideologically defined support groups:

- `sidea_religious_support`
- `sidea_party_elite_support`
- `sidea_rural_worker_support`
- `sidea_military_support`
- `sidea_ethnic_racial_support`

**`legit_ratio`** (H5, H6, H8): The ideological legitimation share, derived in `02_data_prep.R` as:
$$
\text{LegitRatio}_{it} = \frac{v2exl\_legitideol_{it}}{v2exl\_legitideol_{it} + v2exl\_legitperf_{it} + v2exl\_legitlead_{it} + 0.01}
$$
The small constant (0.01) prevents division by zero. Values range from 0 (no ideological legitimation relative to total) to approaching 1 (ideological legitimation dominates).

**`sidea_dynamic_leader`** (H8 moderator): GRAVE-D binary indicator for dynamic or charismatic personal leadership, used to test the Messianic Autocrat hypothesis.

### Control Variables

| Variable | Source | Role |
|---|---|---|
| `log_cinc_a`, `log_cinc_b` | COW NMC | Capability controls (log-transformed) |
| `cinc_ratio` | COW NMC | Relative capability balance |
| `sidea_winning_coalition_size` | V-Dem/BdM | Selectorate theory control |
| `cold_war` | Coded 1947–1991 | Historical period control |
| `t`, `t2`, `t3` | Beck, Katz, Tucker (1998) | Temporal autocorrelation splines |

## Model Specifications

### Tier 1: Binary Time-Series Cross-Sectional Logit

All Tier 1 models are standard logistic regressions (GLM, binomial family). Peace years splines (`t`, `t2`, `t3`) are added to models including temporal controls, following Beck, Katz, and Tucker [-@beck1998] to address temporal autocorrelation in BTSCS data. Each hypothesis is estimated in a sequence of five progressively specified models:

1. IV only (baseline)
2. IV + regime type controls
3. IV + capability controls
4. IV + capability + peace years splines
5. IV + capability + peace years + Cold War dummy (full model)

### Tier 2: Legitimation Mix

H5 and H6 models use `legit_ratio` as the primary IV in the same sequential specification. Hurdle models (`pscl::hurdle`) are estimated alongside standard logits as sensitivity checks for zero-inflation.

### Tier 3: Mediation, Moderation, and Survival

**H7 — Mediation** uses the `mediation` package (Baron-Kenny framework with 1,000 bootstrap simulations) and `lavaan` (SEM). The treatment variable is `sidea_revisionist_domestic`, the mediator is the relevant support group variable, and the outcome is `targets_democracy`.

**H8 — Moderation** estimates interaction models between `legit_ratio` and `sidea_dynamic_leader`. Both logit and linear probability model (LPM) specifications are reported; the LPM facilitates marginal effect interpretation of the interaction term.

**H9 — Survival** uses Cox Proportional Hazard models estimated on the leader-level aggregated dataset. Four models are reported: (1) ideology main effect, (2) legitimation ratio main effect, (3) ideology $\times$ conflict interaction, (4) ratio $\times$ conflict interaction.

## Data Preparation

All data preparation is handled by `R/02_data_prep.R`. Key transformations include:

- **Autocracy filter**: rows with $v2x\_libdem_a \geq 0.5$ are dropped before modeling.
- **Log transformation** of CINC scores: `log_cinc_a = log(cinc_a + 0.001)` to handle near-zero values.
- **Peace years derivation**: the `data.table`-based implementation follows Beck, Katz, and Tucker (1998), with left-censoring resolved by assuming 35 years of peace for regimes with no prior recorded conflict.
- **`legit_ratio` construction**: as specified above.
- **`ideology_gap`**: absolute difference in `v2exl_legitideol` between Side A and Side B, used as a dyadic IV in H3/H4 Model 6 specifications.
- **Output**: cleaned datasets are saved as `.rds` objects in `ready_data/` (Git-ignored) to preserve factor levels and R attributes.


### H8 — Moderation by Dynamic Leadership changes

H8 tests whether dynamic personal leadership qualities moderate the effect of revisionist ideology on conflict behavior. Models estimate the interaction between `legit_ratio` and `sidea_dynamic_leader` (centered to reduce collinearity) on MID initiation and democracy targeting.

Due to severe quasi-complete separation and convergence issues in the full specification (driven by rare events, interactions, and the temporal polynomial), the models were simplified:

- Temporal dependence terms (`t_scaled`, `t2_scaled`, `t3_scaled`) were removed.
- Analysis restricted to Cold War years (1960–1989) to eliminate the need for a `cold_war` dummy.
- Logged CINC variables (`log_cinc_a`, `log_cinc_b`) were dropped.
- Only the core interaction and winning coalition size were retained.

Both logit (cloglog link for rare onset) and linear probability model (LPM) specifications are reported; LPM aids interpretation of interaction marginal effects.

Future extensions (planned): separate runs for the post-Cold War (1990–2000) and post-2001 War on Terror eras to test stability across geopolitical contexts.
# Appendix: Robustness Checks and Descriptive Statistics {#sec-appendix}

```{r}
#| label: setup
#| include: false

library(here)
here::i_am("docs/appendix.qmd")  # Ensure correct project root

source(here::here("R", "00_packages.R"))

```

```{r}


# Load cached dyad_ready for descriptive statistics
dyad_ready_path <- here::here("results", "dyad_ready.rds")
if (file.exists(dyad_ready_path)) {
  dyad_ready <- readRDS(dyad_ready_path)
} else {
  message("dyad_ready.rds not found. Run 02_data_prep.R first.")
}
# Load cached monadic_ready for leader-level descriptive stats
monadic_ready_path <- here::here("results", "monadic_ready.rds")
if (file.exists(monadic_ready_path)) {
  monadic_ready <- readRDS(monadic_ready_path)
} else {
  message("monadic_ready.rds not found. Run 02_data_prep.R first.")
}

```

```{r}

# Load all model objects
h1_models   <- readRDS(here::here("results", "h1_logit_models.rds"))
h2_models   <- readRDS(here::here("results", "h2_logit_models.rds"))
h3_models   <- readRDS(here::here("results", "h3_logit_models.rds"))
h4_models   <- readRDS(here::here("results", "h4_logit_models.rds"))
h5_logit_models <- readRDS(here::here("results", "h5_logit_models.rds"))
h6_logit_models <- readRDS(here::here("results", "h6_logit_models.rds"))
h8_init     <- readRDS(here::here("results", "h8_init_models.rds"))
h8_target   <- readRDS(here::here("results", "h8_target_models.rds"))
# Load H8 robustness models (optional - skip if file missing)
h8_robust <- tryCatch(
  readRDS(here::here("results", "h8_robust_models.rds")),
  error = function(e) {
    message("h8_robust_models.rds not found - skipping H8 robustness in appendix")
    NULL
  }
)
h9_survival <- readRDS(here::here("results", "h9_survival.rds"))

```


```{r}

# Universal coefficient extractor for any model type
# Avoids summary() which crashes on brglmFit and incomplete lm objects
extract_coefs <- function(model, model_name) {
  if (is.null(model)) return(NULL)
  tryCatch({
    if (inherits(model, "logistf")) {
      data.frame(
        model = model_name,
        term = names(coef(model)),
        estimate = coef(model),
        std.error = sqrt(diag(vcov(model))),
        p.value = model$prob,
        stringsAsFactors = FALSE, row.names = NULL
      )
    } else if (inherits(model, "coxph")) {
      s <- summary(model)
      data.frame(
        model = model_name,
        term = rownames(s$coefficients),
        estimate = s$coefficients[, "coef"],
        std.error = s$coefficients[, "se(coef)"],
        p.value = s$coefficients[, "Pr(>|z|)"],
        stringsAsFactors = FALSE, row.names = NULL
      )
    } else {
      # glm, brglmFit, lm - use coef()/vcov() directly
      beta <- coef(model)
      v <- tryCatch(vcov(model), error = function(e) NULL)
      if (is.null(v)) {
        # If vcov fails, try to get SEs from model$coefficients matrix
        se <- rep(NA_real_, length(beta))
      } else {
        se <- sqrt(diag(v))
      }
      z <- beta / se
      pval <- 2 * pnorm(abs(z), lower.tail = FALSE)
      data.frame(
        model = model_name,
        term = names(beta),
        estimate = beta,
        std.error = se,
        p.value = pval,
        stringsAsFactors = FALSE, row.names = NULL
      )
    }
  }, error = function(e) {
    message("Could not extract from ", model_name, ": ", e$message)
    NULL
  })
}

```


```{r}

# Helper to build a kable comparison table from a named list of models
build_model_table <- function(model_list, coef_labels = NULL) {
  all_coefs <- do.call(rbind, Filter(Negate(is.null),
    Map(extract_coefs, model_list, names(model_list))
  ))
  if (is.null(all_coefs) || nrow(all_coefs) == 0) {
    message("No model coefficients could be extracted.")
    return(invisible(NULL))
  }
  if (!is.null(coef_labels)) {
    all_coefs <- all_coefs[all_coefs$term %in% names(coef_labels), ]
    all_coefs$label <- coef_labels[all_coefs$term]
  } else {
    all_coefs$label <- all_coefs$term
  }
  all_coefs$stars <- ifelse(all_coefs$p.value < 0.001, "***",
    ifelse(all_coefs$p.value < 0.01, "**",
      ifelse(all_coefs$p.value < 0.05, "*", "")))
  all_coefs$cell <- paste0(sprintf("%.3f", all_coefs$estimate),
    all_coefs$stars, " (", sprintf("%.3f", all_coefs$std.error), ")")
  wide <- reshape(all_coefs[, c("label", "model", "cell")],
    idvar = "label", timevar = "model", direction = "wide")
  names(wide) <- gsub("cell\\.", "", names(wide))
  knitr::kable(wide, row.names = FALSE)
}
```
This appendix presents supplementary material, including descriptive statistics for all key variables, robustness checks with alternative model specifications, and additional interaction plots.

## Descriptive Statistics

```{r}
#| label: tbl-descriptive
#| tbl-cap: "Descriptive Statistics for Key Variables (Full Dyadic Sample)"
key_vars <- dyad_ready %>%
  dplyr::select(
    mid_initiated,
    targets_democracy,
    sidea_revisionist_domestic,
    legit_ratio,
    v2exl_legitideol_a,
    v2exl_legitperf_a,
    v2exl_legitlead_a,
    sidea_winning_coalition_size,
    cinc_a,
    sidea_religious_support,
    sidea_party_elite_support,
    sidea_military_support,
    sidea_dynamic_leader,
    cold_war
  )
datasummary_skim(
  key_vars,
  type = "numeric",
  title = "Descriptive Statistics: Key Dyadic Variables",
  notes = "Autocratic Side A dyad-years, 1946\u20132010."
)
```

```{r}
#| label: tbl-descriptive-mono
#| tbl-cap: "Descriptive Statistics: Leader-Level Variables (Monadic Sample)"
if (exists("monadic_ready")) {
  mono_vars <- monadic_ready %>%
    dplyr::select(
      tenure_years,
      survived_to_end,
      avg_ideological_legit,
      legit_ratio,
      conflicts_vs_ideo_targets,
      conflicts_vs_democracy,
      avg_autocracy_level,
      log_avg_gdp,
      irregular_entry
    )
  datasummary_skim(
    mono_vars,
    type = "numeric",
    title = "Descriptive Statistics: Leader-Level (Monadic) Variables",
    notes = "Autocratic leader-years, 1946\u20132010."
  )
}
```

## A.2 Correlation Matrix

```{r}
#| label: fig-corr
#| fig-cap: "Correlation matrix of key explanatory variables"
library(ggcorrplot)
corr_data <- dyad_ready %>%
  dplyr::select(
    sidea_revisionist_domestic,
    legit_ratio,
    v2exl_legitperf_a,
    v2exl_legitlead_a,
    sidea_winning_coalition_size,
    sidea_religious_support,
    sidea_military_support,
    sidea_dynamic_leader
  ) %>%
  drop_na()

corr_matrix <- cor(corr_data)

ggcorrplot(
  corr_matrix,
  hc.order = TRUE,
  type = "lower",
  lab = TRUE,
  lab_size = 3,
  ggtheme = ggplot2::theme_minimal()
)
```
