# Data and Methods {#sec-data-methods}

```{r}
#| label: setup
#| include: false
source(here::here("R", "00_packages.R"))
# Load cached dyad_ready (saved by 02_data_prep.R) instead of full pipeline
dyad_ready_path <- here::here("results", "dyad_ready.rds")
if (file.exists(dyad_ready_path)) {
  dyad_ready <- readRDS(dyad_ready_path)
} else {
  message("dyad_ready.rds not found. Run 02_data_prep.R first to generate it.")
}
```

## Data Sources

The primary dataset is **GRAVE_D_Master_with_Leaders.csv**, produced by the GRAVE-D data pipeline (Hanna 2026). GRAVE-D is not itself a primary data collection effort; rather, it is an integration, harmonization, and extension layer built on top of more than a dozen established political science and economics datasets. Its principal contributions are:

- **Structural organization**: converting monadic country-year source data into a directed dyad-year panel using the Formal Bilateral Influence Capacity (FBIC) index [@MoyerEtAl_FBIC] as the dyadic spine, enabling analysis of bilateral relationships;
- **Derived variables**: constructing composite indices (e.g., *revisionist potential*, *ideological extremity*, *religious distance*) and interaction terms not present in any single source;
- **Missingness management**: applying a multi-stage imputation protocol (linear interpolation, regional-year median fill, and derived variable recomputation) documented in the GRAVE-D codebook;
- **Original coding**: ALBA (Bolivarian Alliance) membership indicators.

The underlying source datasets, with their required citations, are:

1. **Formal Bilateral Influence Capacity (FBIC) Index** [@MoyerEtAl_FBIC]: Provides the directed dyad-year spine and all bandwidth variables (total, economic, political, security).

2. **Varieties of Democracy (V-Dem)** v15 [@VDem_Dataset]: Supplies democracy and regime indices, legitimation strategy variables (`v2exl_legitideol`, `v2exl_legitlead`, `v2exl_legitperf`, `v2exl_legitratio`), civil society, corruption, rule of law, power distribution, regime support group, and head-of-state/government variables for both sides of each dyad.

3. **V-Dem Episodes of Regime Transformation (ERT)** [@EdgellEtAl2020_ERT]: Regime transition episodes data.

4. **Correlates of War --- Militarized Interstate Disputes (MIDs)** v5 [@COW_MIDs_v5]: Provides hostility-level data from which `mid_initiated` and conflict outcomes are derived.

5. **Correlates of War --- National Material Capabilities (NMC)** v6 [@COW_NMC_v6]: Provides Composite Index of National Capability (CINC) scores used as controls for material power.

6. **Alliance Treaty Obligations and Provisions (ATOP)** v5.1 [@LeedsEtAl2002_ATOP; @ATOP_Dataset]: Alliance type indicators (defense, offense, neutrality, non-aggression, consultation).

7. **Correlates of War --- World Religion Project** [@COW_WorldReligion]: Religious demographic shares used to compute religious distance variables.

8. **Archigos** [@GoemansEtAl2009_Archigos; @Archigos_Dataset]: Leader tenure, entry/exit modes, and post-tenure fate variables.

9. **Colgan Revolutionary Leaders Dataset** [@Colgan2013_RevolutionaryLeaders; @Colgan_RevolutionaryLeaders_Dataset]: Revolutionary leader classification and revolutionary change category variables.

10. **Global Leader Ideology Dataset** [@Herre2022_GlobalLeaderIdeology_Dataset]: Leader ideology scores, legal tradition, and polity-level controls.

11. **Polity5**: Polity democracy/autocracy scores and regime durability.

12. **First Use of Violent Force Dataset** [@CaprioliTrumbore2003_FirstUse]: First-use-of-force coding for MIDs.

13. **Hanson and Sigman State Capacity Index** [@HansonSigman2021_StateCapacity]: State fiscal and administrative capacity measures.

14. **Fraser Institute Economic Freedom Dataset** [@GwartneyEtAl2022_EFW_Dataset]: Economic freedom indices.

15. **Maddison Project Database** [@BoltVanZanden2014_MaddisonProject; @MaddisonProject_Database]: Historical GDP per capita.

16. **Ross Oil and Gas Dataset** [@Ross_OilGas_Dataset]: Fuel and oil income per capita.

17. **Standardized World Income Inequality Database (SWIID)** [@Solt2020_SWIID; @SWIID_Dataset]: Income inequality (Gini) estimates.

18. **RED --- Relational Export Dataset** [@RED_Dataset]: Trade relationship data.

19. **ALBA Membership** (Hanna 2026): Original coding of Bolivarian Alliance membership.

The sample is restricted to dyad-years in which Side A is classified as autocratic ($v2x\_libdem_a < 0.5$) and the year is 1946 or later (post-World War II international order).

## Variables

### Dependent Variables

**`mid_initiated`** (Tier 1–2 analyses): Binary indicator equal to 1 when Side A initiates a Militarized Interstate Dispute, defined as `hihosta >= 2` (threat or use of force). Constructed in `02_data_prep.R`.

**`targets_democracy`** (Tier 1–2 analyses): Binary indicator equal to 1 when the dyad's Side B is classified as democratic ($v2x\_libdem_b \geq 0.5$), conditional on MID initiation. For H2 and H4 models this variable is the dependent variable on the subset of initiator-years.

**Leader tenure / exit hazard** (Tier 3 survival analysis): Measured in leader-years, with exit coded as an event indicator for Cox Proportional Hazard models.

### Primary Independent Variables

**`sidea_revisionist_domestic`** (H1, H2): GRAVE-D composite measure of leadership revisionist ideology for Side A's governing leadership. Higher values indicate stronger revisionist ideological orientation.

**Support group variables** (H3, H4): GRAVE-D indicators for the regime's reliance on specific ideologically defined support groups:

- `sidea_religious_support`
- `sidea_party_elite_support`
- `sidea_rural_worker_support`
- `sidea_military_support`
- `sidea_ethnic_racial_support`

**`legit_ratio`** (H5, H6, H8): The ideological legitimation share, derived in `02_data_prep.R` as:
$$
\text{LegitRatio}_{it} = \frac{v2exl\_legitideol_{it}}{v2exl\_legitideol_{it} + v2exl\_legitperf_{it} + v2exl\_legitlead_{it} + 0.01}
$$
The small constant (0.01) prevents division by zero. Values range from 0 (no ideological legitimation relative to total) to approaching 1 (ideological legitimation dominates).

**`sidea_dynamic_leader`** (H8 moderator): GRAVE-D binary indicator for dynamic or charismatic personal leadership, used to test the Messianic Autocrat hypothesis.

### Control Variables

| Variable | Source | Role |
|---|---|---|
| `log_cinc_a`, `log_cinc_b` | COW NMC | Capability controls (log-transformed) |
| `cinc_ratio` | COW NMC | Relative capability balance |
| `sidea_winning_coalition_size` | V-Dem/BdM | Selectorate theory control |
| `cold_war` | Coded 1947–1991 | Historical period control |
| `t`, `t2`, `t3` | Beck, Katz, Tucker (1998) | Temporal autocorrelation splines |

## Model Specifications

### Tier 1: Binary Time-Series Cross-Sectional Logit

All Tier 1 models are standard logistic regressions (GLM, binomial family). Peace years splines (`t`, `t2`, `t3`) are added to models including temporal controls, following Beck, Katz, and Tucker [-@beck1998] to address temporal autocorrelation in BTSCS data. Each hypothesis is estimated in a sequence of five progressively specified models:

1. IV only (baseline)
2. IV + regime type controls
3. IV + capability controls
4. IV + capability + peace years splines
5. IV + capability + peace years + Cold War dummy (full model)

### Tier 2: Legitimation Mix

H5 and H6 models use `legit_ratio` as the primary IV in the same sequential specification. Hurdle models (`pscl::hurdle`) are estimated alongside standard logits as sensitivity checks for zero-inflation.

### Tier 3: Mediation, Moderation, and Survival

**H7 — Mediation** uses the `mediation` package (Baron-Kenny framework with 1,000 bootstrap simulations) and `lavaan` (SEM). The treatment variable is `sidea_revisionist_domestic`, the mediator is the relevant support group variable, and the outcome is `targets_democracy`.

**H8 — Moderation** estimates interaction models between `legit_ratio` and `sidea_dynamic_leader`. Both logit and linear probability model (LPM) specifications are reported; the LPM facilitates marginal effect interpretation of the interaction term.

**H9 — Survival** uses Cox Proportional Hazard models estimated on the leader-level aggregated dataset. Four models are reported: (1) ideology main effect, (2) legitimation ratio main effect, (3) ideology $\times$ conflict interaction, (4) ratio $\times$ conflict interaction.

## Data Preparation

All data preparation is handled by `R/02_data_prep.R`. Key transformations include:

- **Autocracy filter**: rows with $v2x\_libdem_a \geq 0.5$ are dropped before modeling.
- **Log transformation** of CINC scores: `log_cinc_a = log(cinc_a + 0.001)` to handle near-zero values.
- **Peace years derivation**: the `data.table`-based implementation follows Beck, Katz, and Tucker (1998), with left-censoring resolved by assuming 35 years of peace for regimes with no prior recorded conflict.
- **`legit_ratio` construction**: as specified above.
- **`ideology_gap`**: absolute difference in `v2exl_legitideol` between Side A and Side B, used as a dyadic IV in H3/H4 Model 6 specifications.
- **Output**: cleaned datasets are saved as `.rds` objects in `ready_data/` (Git-ignored) to preserve factor levels and R attributes.

```{r}
#| label: descriptive-stats
#| tbl-cap: "Descriptive statistics for key analysis variables"

dyad_ready %>%
    dplyr::select(
    mid_initiated, targets_democracy,
    sidea_revisionist_domestic, legit_ratio,
    sidea_dynamic_leader, log_cinc_a, log_cinc_b,
    sidea_winning_coalition_size, cold_war
  ) %>%
  modelsummary::datasummary_skim(
    fmt = 2,
    histogram = FALSE,
    title = "Descriptive Statistics"
  )
```
