# Data and Methods {#sec-data-methods}

```{r}
#| label: setup
#| include: false
source(here::here("R", "00_packages.R"))
source(here::here("R", "01_load_data.R"))
source(here::here("R", "02_data_prep.R"))
```

## Data Sources

The primary dataset is **GRAVE_D_Master_with_Leaders.csv**, which integrates four source datasets:

1. **GRAVE-D** (Geopolitical Revisionism and Violent Extremism Dataset — Dyadic): Provides dyadic militarized interstate dispute data with GRAVE-D leadership ideology scores and support group composition variables for each state's governing regime on the conflict-initiating side.

2. **V-Dem** (Varieties of Democracy, v13): Supplies the legitimation strategy variables used in Tier 2 analyses — `v2exl_legitideol` (ideological legitimation), `v2exl_legitlead` (personalist/leader legitimation), and `v2exl_legitperf` (performance legitimation) — as well as the liberal democracy index used to classify autocracies.

3. **Correlates of War — National Material Capabilities** (v6.0): Provides the Composite Index of National Capability (CINC) scores used as controls for material power balance.

4. **Leader tenure and entry/exit data**: Merged from the Archigos dataset [@goemans2009archigos] via the GRAVE-D master file to support leader-level survival analysis.

The sample is restricted to dyad-years in which Side A is classified as autocratic ($v2x\_libdem_a < 0.5$) and the year is 1946 or later (post-World War II international order).

## Variables

### Dependent Variables

**`mid_initiated`** (Tier 1–2 analyses): Binary indicator equal to 1 when Side A initiates a Militarized Interstate Dispute, defined as `hihosta >= 2` (threat or use of force). Constructed in `02_data_prep.R`.

**`targets_democracy`** (Tier 1–2 analyses): Binary indicator equal to 1 when the dyad's Side B is classified as democratic ($v2x\_libdem_b \geq 0.5$), conditional on MID initiation. For H2 and H4 models this variable is the dependent variable on the subset of initiator-years.

**Leader tenure / exit hazard** (Tier 3 survival analysis): Measured in leader-years, with exit coded as an event indicator for Cox Proportional Hazard models.

### Primary Independent Variables

**`sidea_revisionist_domestic`** (H1, H2): GRAVE-D composite measure of leadership revisionist ideology for Side A's governing leadership. Higher values indicate stronger revisionist ideological orientation.

**Support group variables** (H3, H4): GRAVE-D indicators for the regime's reliance on specific ideologically defined support groups:

- `sidea_religious_support`
- `sidea_party_elite_support`
- `sidea_rural_worker_support`
- `sidea_military_support`
- `sidea_ethnic_racial_support`

**`legit_ratio`** (H5, H6, H8): The ideological legitimation share, derived in `02_data_prep.R` as:
$$
\text{LegitRatio}_{it} = \frac{v2exl\_legitideol_{it}}{v2exl\_legitideol_{it} + v2exl\_legitperf_{it} + v2exl\_legitlead_{it} + 0.01}
$$
The small constant (0.01) prevents division by zero. Values range from 0 (no ideological legitimation relative to total) to approaching 1 (ideological legitimation dominates).

**`sidea_dynamic_leader`** (H8 moderator): GRAVE-D binary indicator for dynamic or charismatic personal leadership, used to test the Messianic Autocrat hypothesis.

### Control Variables

| Variable | Source | Role |
|---|---|---|
| `log_cinc_a`, `log_cinc_b` | COW NMC | Capability controls (log-transformed) |
| `cinc_ratio` | COW NMC | Relative capability balance |
| `sidea_winning_coalition_size` | V-Dem/BdM | Selectorate theory control |
| `cold_war` | Coded 1947–1991 | Historical period control |
| `t`, `t2`, `t3` | Beck, Katz, Tucker (1998) | Temporal autocorrelation splines |

## Model Specifications

### Tier 1: Binary Time-Series Cross-Sectional Logit

All Tier 1 models are standard logistic regressions (GLM, binomial family). Peace years splines (`t`, `t2`, `t3`) are added to models including temporal controls, following Beck, Katz, and Tucker [-@beck1998] to address temporal autocorrelation in BTSCS data. Each hypothesis is estimated in a sequence of five progressively specified models:

1. IV only (baseline)
2. IV + regime type controls
3. IV + capability controls
4. IV + capability + peace years splines
5. IV + capability + peace years + Cold War dummy (full model)

### Tier 2: Legitimation Mix

H5 and H6 models use `legit_ratio` as the primary IV in the same sequential specification. Hurdle models (`pscl::hurdle`) are estimated alongside standard logits as sensitivity checks for zero-inflation.

### Tier 3: Mediation, Moderation, and Survival

**H7 — Mediation** uses the `mediation` package (Baron-Kenny framework with 1,000 bootstrap simulations) and `lavaan` (SEM). The treatment variable is `sidea_revisionist_domestic`, the mediator is the relevant support group variable, and the outcome is `targets_democracy`.

**H8 — Moderation** estimates interaction models between `legit_ratio` and `sidea_dynamic_leader`. Both logit and linear probability model (LPM) specifications are reported; the LPM facilitates marginal effect interpretation of the interaction term.

**H9 — Survival** uses Cox Proportional Hazard models estimated on the leader-level aggregated dataset. Four models are reported: (1) ideology main effect, (2) legitimation ratio main effect, (3) ideology $\times$ conflict interaction, (4) ratio $\times$ conflict interaction.

## Data Preparation

All data preparation is handled by `R/02_data_prep.R`. Key transformations include:

- **Autocracy filter**: rows with $v2x\_libdem_a \geq 0.5$ are dropped before modeling.
- **Log transformation** of CINC scores: `log_cinc_a = log(cinc_a + 0.001)` to handle near-zero values.
- **Peace years derivation**: the `data.table`-based implementation follows Beck, Katz, and Tucker (1998), with left-censoring resolved by assuming 35 years of peace for regimes with no prior recorded conflict.
- **`legit_ratio` construction**: as specified above.
- **`ideology_gap`**: absolute difference in `v2exl_legitideol` between Side A and Side B, used as a dyadic IV in H3/H4 Model 6 specifications.
- **Output**: cleaned datasets are saved as `.rds` objects in `ready_data/` (Git-ignored) to preserve factor levels and R attributes.

```{r}
#| label: descriptive-stats
#| tbl-cap: "Descriptive statistics for key analysis variables"

dyad_ready %>%
  select(
    mid_initiated, targets_democracy,
    sidea_revisionist_domestic, legit_ratio,
    sidea_dynamic_leader, log_cinc_a, log_cinc_b,
    sidea_winning_coalition_size, cold_war
  ) %>%
  modelsummary::datasummary_skim(
    fmt = 2,
    histogram = FALSE,
    title = "Descriptive Statistics"
  )
```
