# Tier 2 Results: Legitimation Mix (H5–H6) {#sec-results-tier2}

```{r}
#| label: setup
#| include: false
source(here::here("R", "00_packages.R"))

# Load pre-estimated models
h5_logit_models <- readRDS(here::here("results", "h5_logit_models.rds"))
h5_hurdle_models <- readRDS(here::here("results", "h5_hurdle_models.rds"))

# Load cached dyad_ready for prediction plots
dyad_ready_path <- here::here("results", "dyad_ready.rds")
if (file.exists(dyad_ready_path)) {
  dyad_ready <- readRDS(dyad_ready_path)
} else {
  message("dyad_ready.rds not found. Prediction plots will fail.")
}

h6_logit_models <- readRDS(here::here("results", "h6_logit_models.rds"))

# Universal coefficient extractor (avoids summary() which crashes on brglmFit)
extract_coefs <- function(model, model_name) {
  if (is.null(model)) return(NULL)
  tryCatch({
    if (inherits(model, "logistf")) {
      data.frame(model = model_name, term = names(coef(model)),
                 estimate = coef(model), std.error = sqrt(diag(vcov(model))),
                 p.value = model$prob, stringsAsFactors = FALSE, row.names = NULL)
    } else if (inherits(model, "coxph")) {
      s <- summary(model)
      data.frame(model = model_name, term = rownames(s$coefficients),
                 estimate = s$coefficients[, "coef"], std.error = s$coefficients[, "se(coef)"],
                 p.value = s$coefficients[, "Pr(>|z|)"], stringsAsFactors = FALSE, row.names = NULL)
    } else {
      beta <- coef(model)
      v <- tryCatch(vcov(model), error = function(e) NULL)
      se <- if (is.null(v)) rep(NA_real_, length(beta)) else sqrt(diag(v))
      z <- beta / se
      pval <- 2 * pnorm(abs(z), lower.tail = FALSE)
      data.frame(model = model_name, term = names(beta),
                 estimate = beta, std.error = se, p.value = pval,
                 stringsAsFactors = FALSE, row.names = NULL)
    }
  }, error = function(e) { message("Could not extract from ", model_name, ": ", e$message); NULL })
}

build_model_table <- function(model_list, coef_labels = NULL) {
  all_coefs <- do.call(rbind, Filter(Negate(is.null),
    Map(extract_coefs, model_list, names(model_list))))
  if (is.null(all_coefs) || nrow(all_coefs) == 0) {
    message("No model coefficients could be extracted.")
    return(invisible(NULL))
  }
  if (!is.null(coef_labels)) {
    all_coefs <- all_coefs[all_coefs$term %in% names(coef_labels), ]
    all_coefs$label <- coef_labels[all_coefs$term]
  } else {
    all_coefs$label <- all_coefs$term
  }
  all_coefs$stars <- ifelse(all_coefs$p.value < 0.001, "***",
    ifelse(all_coefs$p.value < 0.01, "**",
    ifelse(all_coefs$p.value < 0.05, "*", "")))
  all_coefs$cell <- paste0(sprintf("%.3f", all_coefs$estimate),
    all_coefs$stars, " (", sprintf("%.3f", all_coefs$std.error), ")")
  wide <- reshape(all_coefs[, c("label", "model", "cell")],
    idvar = "label", timevar = "model", direction = "wide")
  names(wide) <- gsub("cell\\.", "", names(wide))
  knitr::kable(wide, row.names = FALSE)
}
```

This chapter presents results for Tier 2 hypotheses, which examine how the relative dependence on ideological legitimation—measured as a ratio of ideological to total legitimation score (`legit_ratio`)—shapes both conflict initiation and target selection. Models include standard logistic regressions and two-part hurdle models. All models are estimated on the full dyadic dataset (`dyad_ready`) with Side A restricted to autocracies ($v2x\_libdem_a < 0.5$).

## H5: Legitimation Mix and Conflict Initiation

**H5** predicts that autocratic states with greater relative dependence on ideological legitimation will be more likely to initiate MIDs. The dependent variable is `mid_initiated` (hihosta $\geq$ 2). Four sequentially specified logit models are estimated, followed by a two-part hurdle specification (@tbl-h5-logit, @tbl-h5-hurdle).

```{r}
#| label: tbl-h5-logit
#| tbl-cap: "H5: Legitimation Mix and Conflict Initiation (Logistic Regression)"
build_model_table(
  list(
    "(1) Baseline" = h5_logit_models$h5_baseline,
    "(2) +Components" = h5_logit_models$h5_components,
    "(3) +Controls" = h5_logit_models$h5_controls,
    "(4) Full" = h5_logit_models$h5_full
  )
)
```

```{r}
#| label: tbl-h5-hurdle
#| tbl-cap: "H5: Hurdle Model — Binary and Count Parts"
build_model_table(
  list(
    "Part 1: Binary" = h5_hurdle_models$hurdle_binary,
    "Part 2: Count" = h5_hurdle_models$hurdle_count
  )
)
```

```{r}
#| label: fig-h5-pred
#| fig-cap: "H5: Predicted probability of MID initiation by legitimation ratio (full logit model)"

# Predicted probability plot for H5 full model
pred_h5 <- data.frame(
  legit_ratio = seq(
    quantile(dyad_ready$legit_ratio, 0.05, na.rm = TRUE),
    quantile(dyad_ready$legit_ratio, 0.95, na.rm = TRUE),
    length.out = 100
  )
) %>%
  mutate(
    v2exl_legitperf_a = mean(dyad_ready$v2exl_legitperf_a, na.rm = TRUE),
    v2exl_legitlead_a = mean(dyad_ready$v2exl_legitlead_a, na.rm = TRUE),
    targets_democracy = 0,
    cinc_a = mean(
      dyad_ready$cinc_a, na.rm = TRUE),
    sidea_winning_coalition_size = mean(
      dyad_ready$sidea_winning_coalition_size, na.rm = TRUE),
    cold_war = 0,
    t = 0, t2 = 0, t3 = 0
  )

pred_h5 <- pred_h5 %>%
  bind_cols(
    predict(h5_logit_models$h5_full, newdata = pred_h5,
            type = "link", se.fit = TRUE) %>%
      as.data.frame() %>%
      transmute(
        fit = plogis(fit),
        lower = plogis(fit - 1.96 * se.fit),
        upper = plogis(fit + 1.96 * se.fit)
      )
  )

ggplot(pred_h5, aes(x = legit_ratio, y = fit)) +
  geom_line(color = "#0072B2", linewidth = 1.2) +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2, fill = "#0072B2") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 0.1)) +
  labs(
    x = "Ideological Legitimation Ratio (V-Dem)",
    y = "Predicted Probability of MID Initiation",
    subtitle = "Other variables held at means; Cold War = 0; Targets Democracy = 0"
  )
```

## H6: Legitimation Mix and Democracy Targeting

**H6** predicts that among conflict initiators, states with higher relative ideological legitimation preferentially target democracies. The sample is restricted to `mid_initiated == 1` rows; the dependent variable is `targets_democracy`. An additional interaction model tests whether the legitimation mix effect is amplified by leader revisionist ideology (@tbl-h6).

```{r}
#| label: tbl-h6
#| tbl-cap: "H6: Legitimation Mix and Democracy Targeting (Initiators Only)"
build_model_table(
  list(
    "(1) Baseline" = h6_logit_models$h6_baseline,
    "(2) +Components" = h6_logit_models$h6_components,
    "(3) +Controls" = h6_logit_models$h6_controls,
    "(4) Full" = h6_logit_models$h6_full,
    "(5) Interaction" = h6_logit_models$h6_interaction
  )
)
```

```{r}
#| label: fig-h6-pred
#| fig-cap: "H6: Predicted probability of democracy targeting by legitimation ratio (full model)"

# Load conflict-only data for prediction
conflict_data <- dyad_ready %>% filter(mid_initiated == 1)

pred_h6 <- data.frame(
  legit_ratio = seq(
    quantile(conflict_data$legit_ratio, 0.05, na.rm = TRUE),
    quantile(conflict_data$legit_ratio, 0.95, na.rm = TRUE),
    length.out = 100
  )
) %>%
  mutate(
    v2exl_legitperf_a = mean(conflict_data$v2exl_legitperf_a, na.rm = TRUE),
    v2exl_legitlead_a = mean(conflict_data$v2exl_legitlead_a, na.rm = TRUE),
    cinc_a = mean(
      conflict_data$cinc_a, na.rm = TRUE),
    sidea_winning_coalition_size = mean(
      conflict_data$sidea_winning_coalition_size, na.rm = TRUE),
    cold_war = 0,
    t = 0
  )

pred_h6 <- pred_h6 %>%
  bind_cols(
    predict(h6_logit_models$h6_full, newdata = pred_h6,
            type = "link", se.fit = TRUE) %>%
      as.data.frame() %>%
      transmute(
        fit = plogis(fit),
        lower = plogis(fit - 1.96 * se.fit),
        upper = plogis(fit + 1.96 * se.fit)
      )
  )

ggplot(pred_h6, aes(x = legit_ratio, y = fit)) +
  geom_line(color = "#009E73", linewidth = 1.2) +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2, fill = "#009E73") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 0.1)) +
  labs(
    x = "Ideological Legitimation Ratio (V-Dem)",
    y = "Predicted Probability of Targeting a Democracy",
    subtitle = "Other variables held at means; Cold War = 0"
  )
```

## Summary

@tbl-tier2-summary provides a compact side-by-side summary of the key legitimation ratio coefficients across H5 and H6 full models.

```{r}
#| label: tbl-tier2-summary
#| tbl-cap: "Tier 2 Summary: Legitimation Ratio Estimates (H5–H6 Full Models)"
build_model_table(
  list(
    "H5: Initiation" = h5_logit_models$h5_full,
    "H6: Targeting" = h6_logit_models$h6_full,
    "H6: Interaction" = h6_logit_models$h6_interaction
  )
)
```
