# Tier 2 Results: Legitimation Mix (H5–H6) {#sec-results-tier2}

```{r}
#| label: setup
#| include: false
source(here::here("R", "00_packages.R"))

# Load pre-estimated models
h5_logit_models <- readRDS(here::here("results", "h5_logit_models.rds"))
h5_hurdle_models <- readRDS(here::here("results", "h5_hurdle_models.rds"))

# Load cached dyad_ready for prediction plots
dyad_ready_path <- here::here("results", "dyad_ready.rds")
if (file.exists(dyad_ready_path)) {
  dyad_ready <- readRDS(dyad_ready_path)
} else {
  message("dyad_ready.rds not found. Prediction plots will fail.")
}
h6_logit_models <- readRDS(here::here("results", "h6_logit_models.rds"))
```

This chapter presents results for Tier 2 hypotheses, which examine how the relative dependence on ideological legitimation—measured as a ratio of ideological to total legitimation score (`legit_ratio`)—shapes both conflict initiation and target selection. Models include standard logistic regressions and two-part hurdle models. All models are estimated on the full dyadic dataset (`dyad_ready`) with Side A restricted to autocracies ($v2x\_libdem_a < 0.5$).

## H5: Legitimation Mix and Conflict Initiation

**H5** predicts that autocratic states with greater relative dependence on ideological legitimation will be more likely to initiate MIDs. The dependent variable is `mid_initiated` (hihosta $\geq$ 2). Four sequentially specified logit models are estimated, followed by a two-part hurdle specification (@tbl-h5-logit, @tbl-h5-hurdle).

```{r}
#| label: tbl-h5-logit
#| tbl-cap: "H5: Legitimation Mix and Conflict Initiation (Logistic Regression)"
#| results: asis
stargazer(
  h5_logit_models$h5_baseline,
  h5_logit_models$h5_components,
  h5_logit_models$h5_controls,
  h5_logit_models$h5_full,
  type = "html",
  title = "H5: Legitimation Mix and Conflict Initiation",
  dep.var.labels = "MID Initiated",
  column.labels = c("(1) Baseline", "(2) +Components", "(3) +Controls", "(4) Full"),
  covariate.labels = c(
    "Legitimation Ratio",
    "Perf. Legitimation", "Personal. Legitimation",
    "Targets Democracy",
    "CINC (A)", "Winning Coalition",
    "Peace Years (t)", "Peace Years (t\\textsuperscript{2})",
    "Peace Years (t\\textsuperscript{3})", "Cold War"
  ),
  star.cutoffs = c(0.05, 0.01, 0.001),
  omit.stat = c("aic", "ll"),
  notes = c("Sample: Autocratic dyad-years, 1946\u20132010.",
            "*** p<0.001, ** p<0.01, * p<0.05")
)
```

```{r}
#| label: tbl-h5-hurdle
#| tbl-cap: "H5: Hurdle Model — Binary and Count Parts"
#| results: asis
stargazer(
  h5_hurdle_models$hurdle_binary,
  h5_hurdle_models$hurdle_count,
  type = "html",
  title = "H5: Two-Part Hurdle Model (Initiation)",
  dep.var.labels = c("Any MID (Logit)", "MID Count (Poisson)"),
  column.labels = c("Part 1: Binary", "Part 2: Count"),
  covariate.labels = c(
    "Legitimation Ratio",
    "Perf. Legitimation", "Personal. Legitimation",
    "Targets Democracy",
    "CINC (A)", "Winning Coalition",
    "Peace Years (t)", "Peace Years (t\\textsuperscript{2})",
    "Peace Years (t\\textsuperscript{3})", "Cold War"
  ),
  star.cutoffs = c(0.05, 0.01, 0.001),
  omit.stat = c("aic", "ll"),
  notes = c("Part 2 estimated on MID initiators only.",
            "*** p<0.001, ** p<0.01, * p<0.05")
)
```

```{r}
#| label: fig-h5-pred
#| fig-cap: "H5: Predicted probability of MID initiation by legitimation ratio (full logit model)"

# Predicted probability plot for H5 full model
pred_h5 <- data.frame(
  legit_ratio = seq(
    quantile(dyad_ready$legit_ratio, 0.05, na.rm = TRUE),
    quantile(dyad_ready$legit_ratio, 0.95, na.rm = TRUE),
    length.out = 100
  )
) %>%
  mutate(
    v2exl_legitperf_a = mean(dyad_ready$v2exl_legitperf_a, na.rm = TRUE),
    v2exl_legitlead_a = mean(dyad_ready$v2exl_legitlead_a, na.rm = TRUE),
    targets_democracy = 0,
    cinc_a = mean(
      dyad_ready$cinc_a, na.rm = TRUE),
    sidea_winning_coalition_size = mean(
      dyad_ready$sidea_winning_coalition_size, na.rm = TRUE),
    cold_war = 0,
    t = 0, t2 = 0, t3 = 0
  )

pred_h5 <- pred_h5 %>%
  bind_cols(
    predict(h5_logit_models$h5_full, newdata = pred_h5,
            type = "link", se.fit = TRUE) %>%
      as.data.frame() %>%
      transmute(
        fit   = plogis(fit),
        lower = plogis(fit - 1.96 * se.fit),
        upper = plogis(fit + 1.96 * se.fit)
      )
  )

ggplot(pred_h5, aes(x = legit_ratio, y = fit)) +
  geom_line(color = "#0072B2", linewidth = 1.2) +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2, fill = "#0072B2") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 0.1)) +
  labs(
    x = "Ideological Legitimation Ratio (V-Dem)",
    y = "Predicted Probability of MID Initiation",
    subtitle = "Other variables held at means; Cold War = 0; Targets Democracy = 0"
  )
```

## H6: Legitimation Mix and Democracy Targeting

**H6** predicts that among conflict initiators, states with higher relative ideological legitimation preferentially target democracies. The sample is restricted to `mid_initiated == 1` rows; the dependent variable is `targets_democracy`. An additional interaction model tests whether the legitimation mix effect is amplified by leader revisionist ideology (@tbl-h6).

```{r}
#| label: tbl-h6
#| tbl-cap: "H6: Legitimation Mix and Democracy Targeting (Initiators Only)"
#| results: asis
stargazer(
  h6_logit_models$h6_baseline,
  h6_logit_models$h6_components,
  h6_logit_models$h6_controls,
  h6_logit_models$h6_full,
  h6_logit_models$h6_interaction,
  type = "html",
  title = "H6: Legitimation Mix and Democracy Targeting",
  dep.var.labels = "Target is Democracy",
  column.labels = c("(1) Baseline", "(2) +Components", "(3) +Controls",
                     "(4) Full", "(5) Interaction"),
  covariate.labels = c(
    "Legitimation Ratio",
    "Perf. Legitimation", "Personal. Legitimation",
    "CINC (A)", "Winning Coalition",
    "Peace Years (t)", "Cold War",
    "Revisionist Ideology",
    "Legit. Ratio × Rev. Ideology"
  ),
  star.cutoffs = c(0.05, 0.01, 0.001),
  omit.stat = c("aic", "ll"),
  notes = c("Sample: Autocratic MID initiators only.",
            "*** p<0.001, ** p<0.01, * p<0.05")
)
```

```{r}
#| label: fig-h6-pred
#| fig-cap: "H6: Predicted probability of democracy targeting by legitimation ratio (full model)"

# Load conflict-only data for prediction
conflict_data <- dyad_ready %>% filter(mid_initiated == 1)

pred_h6 <- data.frame(
  legit_ratio = seq(
    quantile(conflict_data$legit_ratio, 0.05, na.rm = TRUE),
    quantile(conflict_data$legit_ratio, 0.95, na.rm = TRUE),
    length.out = 100
  )
) %>%
  mutate(
    v2exl_legitperf_a = mean(conflict_data$v2exl_legitperf_a, na.rm = TRUE),
    v2exl_legitlead_a = mean(conflict_data$v2exl_legitlead_a, na.rm = TRUE),
    cinc_a = mean(
      conflict_data$cinc_a, na.rm = TRUE),
    sidea_winning_coalition_size = mean(
      conflict_data$sidea_winning_coalition_size, na.rm = TRUE),
    cold_war = 0,
    t = 0
  )

pred_h6 <- pred_h6 %>%
  bind_cols(
    predict(h6_logit_models$h6_full, newdata = pred_h6,
            type = "link", se.fit = TRUE) %>%
      as.data.frame() %>%
      transmute(
        fit   = plogis(fit),
        lower = plogis(fit - 1.96 * se.fit),
        upper = plogis(fit + 1.96 * se.fit)
      )
  )

ggplot(pred_h6, aes(x = legit_ratio, y = fit)) +
  geom_line(color = "#009E73", linewidth = 1.2) +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2, fill = "#009E73") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 0.1)) +
  labs(
    x = "Ideological Legitimation Ratio (V-Dem)",
    y = "Predicted Probability of Targeting a Democracy",
    subtitle = "Other variables held at means; Cold War = 0"
  )
```

## Summary

@tbl-tier2-summary provides a compact side-by-side summary of the key legitimation ratio coefficients across H5 and H6 full models.

```{r}
#| label: tbl-tier2-summary
#| tbl-cap: "Tier 2 Summary: Legitimation Ratio Estimates (H5–H6 Full Models)"
modelsummary(
  list(
    "H5: Initiation"  = h5_logit_models$h5_full,
    "H6: Targeting"    = h6_logit_models$h6_full,
    "H6: Interaction"  = h6_logit_models$h6_interaction
  ),
  coef_map = c(
    legit_ratio = "Legitimation Ratio",
    v2exl_legitperf_a = "Perf. Legitimation",
    v2exl_legitlead_a = "Personal. Legitimation",
    sidea_revisionist_domestic = "Revisionist Ideology",
    `legit_ratio:sidea_revisionist_domestic` = "Legit. Ratio × Rev. Ideology",
    sidea_winning_coalition_size = "Winning Coalition",
    cold_war = "Cold War"
  ),
  stars = c("*" = 0.05, "**" = 0.01, "***" = 0.001),
  gof_omit = "AIC|BIC|Log"
)
```
