# Tier 1 Results: Simple Logistic Regression (H1–H4) {#sec-results-tier1}

```{r}
#| label: setup
#| include: false
source(here::here("R", "00_packages.R"))
source(here::here("R", "02_data_prep.R"))

# Load pre-estimated models
h1_models <- readRDS(here::here("results", "h1_logit_models.rds"))
h2_models <- readRDS(here::here("results", "h2_logit_models.rds"))
h3_models <- readRDS(here::here("results", "h3_logit_models.rds"))
h4_models <- readRDS(here::here("results", "h4_logit_models.rds"))
```

This chapter presents results for Tier 1 hypotheses, which use standard logistic regression to estimate the relationship between leadership ideology, support group composition, and conflict behavior (initiation and target selection). All models are estimated on the full dyadic dataset (`dyad_ready`) with Side A restricted to autocracies ($v2x\_libdem_a < 0.5$).

## H1: Leadership Ideology and Conflict Initiation

**H1** predicts that autocratic states with higher revisionist domestic leadership ideology will be more likely to initiate MIDs. The dependent variable is `mid_initiated` (hihosta $\geq$ 2).

```{r}
#| label: tbl-h1
#| tbl-cap: "H1: Revisionist Leadership Ideology and Conflict Initiation"
#| results: asis
stargazer(
  h1_models$h1_baseline,
  h1_models$h1_controls,
  h1_models$h1_full,
  type = "html",
  title = "H1: Revisionist Leadership Ideology and Conflict Initiation",
  dep.var.labels = "MID Initiated",
  column.labels = c("(1) Baseline", "(2) +Controls", "(3) Full"),
  star.cutoffs = c(0.05, 0.01, 0.001),
  omit.stat = c("aic", "ll"),
  notes = c("Sample: Autocratic dyad-years, 1946\u20132010.",
            "*** p<0.001, ** p<0.01, * p<0.05")
)
```

### Ideology Sub-type Models

```{r}
#| label: tbl-h1-subtypes
#| tbl-cap: "H1: Ideology Sub-types and Conflict Initiation"
#| results: asis
stargazer(
  h1_models$h1_religious,
  h1_models$h1_socialist,
  h1_models$h1_nationalist,
  type = "html",
  title = "H1: Ideology Sub-types and Conflict Initiation",
  dep.var.labels = "MID Initiated",
  column.labels = c("Religious", "Socialist", "Nationalist"),
  star.cutoffs = c(0.05, 0.01, 0.001),
  omit.stat = c("aic", "ll")
)
```

## H2: Leadership Ideology and Democracy Targeting

**H2** predicts that among conflict initiators, states with higher revisionist ideology preferentially target democracies. The sample is restricted to `mid_initiated == 1` rows.

```{r}
#| label: tbl-h2
#| tbl-cap: "H2: Revisionist Leadership Ideology and Democracy Targeting (Initiators Only)"
#| results: asis
stargazer(
  h2_models$h2_baseline,
  h2_models$h2_controls,
  h2_models$h2_full,
  type = "html",
  title = "H2: Leadership Ideology and Democracy Targeting",
  dep.var.labels = "Target is Democracy",
  column.labels = c("(1) Baseline", "(2) +Controls", "(3) Full"),
  star.cutoffs = c(0.05, 0.01, 0.001),
  omit.stat = c("aic", "ll"),
  notes = c("Sample: Autocratic MID initiators only.",
            "*** p<0.001, ** p<0.01, * p<0.05")
)
```

## H3: Support Group Ideology and Conflict Initiation

**H3** tests the Rational Autocrat prediction that regime dependence on ideologically revisionist support groups drives conflict initiation.

```{r}
#| label: tbl-h3
#| tbl-cap: "H3: Support Group Ideology and Conflict Initiation"
#| results: asis
stargazer(
  h3_models$h3_baseline,
  h3_models$h3_party,
  h3_models$h3_military,
  h3_models$h3_multi,
  h3_models$h3_controls,
  h3_models$h3_full,
  type = "html",
  title = "H3: Support Group Ideology and Conflict Initiation",
  dep.var.labels = "MID Initiated",
  column.labels = c("(1) Religious", "(2) Party", "(3) Military",
                    "(4) Multi", "(5) +Controls", "(6) Full"),
  star.cutoffs = c(0.05, 0.01, 0.001),
  omit.stat = c("aic", "ll")
)
```

## H4: Support Group Ideology and Democracy Targeting

**H4** tests whether support group ideology shapes target selection among conflict initiators.

```{r}
#| label: tbl-h4
#| tbl-cap: "H4: Support Group Ideology and Democracy Targeting (Initiators Only)"
#| results: asis
stargazer(
  h4_models$h4_baseline,
  h4_models$h4_party,
  h4_models$h4_military,
  h4_models$h4_multi,
  h4_models$h4_controls,
  h4_models$h4_full,
  type = "html",
  title = "H4: Democracy Targeting (Initiators Only)",
  dep.var.labels = "Target is Democracy",
  column.labels = c("(1) Religious", "(2) Party", "(3) Military",
                    "(4) Multi", "(5) +Controls", "(6) Full"),
  star.cutoffs = c(0.05, 0.01, 0.001),
  omit.stat = c("aic", "ll")
)
```

## Summary

@tbl-tier1-summary provides a compact summary of the key coefficient estimates and significance levels across H1–H4.

```{r}
#| label: tbl-tier1-summary
#| tbl-cap: "Tier 1 Summary: Key Coefficient Estimates (H1\u2013H4 Full Models)"
modelsummary(
  list(
    "H1: Initiation" = h1_models$h1_full,
    "H2: Targeting" = h2_models$h2_full,
    "H3: Initiation" = h3_models$h3_full,
    "H4: Targeting" = h4_models$h4_full
  ),
  stars = c("*" = 0.05, "**" = 0.01, "***" = 0.001),
  gof_omit = "AIC|BIC|Log"
)
```
