# Tier 1 Results: Simple Logistic Regression (H1–H4) {#sec-results-tier1}

```{r}
#| label: setup
#| include: false
source(here::here("R", "00_packages.R"))

# Load pre-estimated models
h1_models <- readRDS(here::here("results", "h1_logit_models.rds"))
h2_models <- readRDS(here::here("results", "h2_logit_models.rds"))
h3_models <- readRDS(here::here("results", "h3_logit_models.rds"))
h4_models <- readRDS(here::here("results", "h4_logit_models.rds"))

# Safe vcov that bypasses summary.brglmFit (which crashes on mu.eta)
safe_vcov <- function(model) {
  if (inherits(model, "brglmFit") && !is.null(model$Fisher.info)) {
    V <- solve(model$Fisher.info)
    nms <- names(coef(model))
    if (nrow(V) == length(nms)) {
      rownames(V) <- colnames(V) <- nms
    }
    return(V)
  }
  if (inherits(model, "glm") && !is.null(model$qr)) {
    p <- model$rank
    p1 <- seq_len(p)
    R <- qr.R(model$qr)[p1, p1, drop = FALSE]
    Rinv <- backsolve(R, diag(p))
    dispersion <- if (model$family$family %in% c("binomial", "poisson")) 1 else {
      sum(model$weights * model$residuals^2) / model$df.residual
    }
    V <- dispersion * tcrossprod(Rinv)
    rownames(V) <- colnames(V) <- names(coef(model))[p1]
    return(V)
  }
  stats::vcov(model)
}

# Universal coefficient extractor (avoids summary() which crashes on brglmFit)
extract_coefs <- function(model, model_name) {
  if (is.null(model)) return(NULL)
  tryCatch({
    if (inherits(model, "logistf")) {
      data.frame(model = model_name, term = names(coef(model)),
        estimate = coef(model), std.error = sqrt(diag(vcov(model))),
        p.value = model$prob, stringsAsFactors = FALSE, row.names = NULL)
    } else if (inherits(model, "coxph")) {
      s <- summary(model)
      data.frame(model = model_name, term = rownames(s$coefficients),
        estimate = s$coefficients[, "coef"], std.error = s$coefficients[, "se(coef)"],
        p.value = s$coefficients[, "Pr(>|z|)"], stringsAsFactors = FALSE, row.names = NULL)
    } else {
      beta <- coef(model)
      v <- tryCatch(safe_vcov(model), error = function(e) NULL)
      se <- if (is.null(v)) rep(NA_real_, length(beta)) else sqrt(diag(v))
      z <- beta / se
      pval <- 2 * pnorm(abs(z), lower.tail = FALSE)
      data.frame(model = model_name, term = names(beta),
        estimate = beta, std.error = se, p.value = pval,
        stringsAsFactors = FALSE, row.names = NULL)
    }
  }, error = function(e) { message("Could not extract from ", model_name, ": ", e$message); NULL })
}

build_model_table <- function(model_list, coef_labels = NULL) {
  all_coefs <- do.call(rbind, Filter(Negate(is.null),
    Map(extract_coefs, model_list, names(model_list))))
  if (is.null(all_coefs) || nrow(all_coefs) == 0) {
    message("No model coefficients could be extracted.")
    return(invisible(NULL))
  }
  if (!is.null(coef_labels)) {
    all_coefs <- all_coefs[all_coefs$term %in% names(coef_labels), ]
    all_coefs$label <- coef_labels[all_coefs$term]
  } else {
    all_coefs$label <- all_coefs$term
  }
  all_coefs$stars <- ifelse(all_coefs$p.value < 0.001, "***",
    ifelse(all_coefs$p.value < 0.01, "**",
    ifelse(all_coefs$p.value < 0.05, "*", "")))
  all_coefs$cell <- paste0(sprintf("%.3f", all_coefs$estimate),
    all_coefs$stars, " (", sprintf("%.3f", all_coefs$std.error), ")")
  wide <- reshape(all_coefs[, c("label", "model", "cell")],
    idvar = "label", timevar = "model", direction = "wide")
  names(wide) <- gsub("cell\\.", "", names(wide))
  knitr::kable(wide, row.names = FALSE)
}
```

This chapter presents results for Tier 1 hypotheses, which use standard logistic regression to estimate the relationship between leadership ideology, support group composition, and conflict behavior (initiation and target selection). All models are estimated on the full dyadic dataset (`dyad_ready`) with Side A restricted to autocracies ($v2x\_libdem_a < 0.5$).

## H1: Leadership Ideology and Conflict Initiation

**H1** predicts that autocratic states with higher revisionist domestic leadership ideology will be more likely to initiate MIDs. The dependent variable is `mid_initiated` (hihosta $\geq$ 2).

```{r}
#| label: tbl-h1
#| tbl-cap: "H1: Revisionist Leadership Ideology and Conflict Initiation"
build_model_table(
  list(
    "(1) Baseline" = h1_models$h1_baseline,
    "(2) +Controls" = h1_models$h1_controls,
    "(3) Full" = h1_models$h1_full
  )
)
```

### Ideology Sub-type Models

```{r}
#| label: tbl-h1-subtypes
#| tbl-cap: "H1: Ideology Sub-types and Conflict Initiation"
build_model_table(
  list(
    "Religious" = h1_models$h1_religious,
    "Socialist" = h1_models$h1_socialist,
    "Nationalist" = h1_models$h1_nationalist
  )
)
```

## H2: Leadership Ideology and Democracy Targeting

**H2** predicts that among conflict initiators, states with higher revisionist ideology preferentially target democracies. The sample is restricted to `mid_initiated == 1` rows.

```{r}
#| label: tbl-h2
#| tbl-cap: "H2: Revisionist Leadership Ideology and Democracy Targeting (Initiators Only)"
build_model_table(
  list(
    "(1) Baseline" = h2_models$h2_baseline,
    "(2) +Controls" = h2_models$h2_controls,
    "(3) Full" = h2_models$h2_full
  )
)
```

## H3: Support Group Ideology and Conflict Initiation

**H3** tests the Rational Autocrat prediction that regime dependence on ideologically revisionist support groups drives conflict initiation.

```{r}
#| label: tbl-h3
#| tbl-cap: "H3: Support Group Ideology and Conflict Initiation"
build_model_table(
  list(
    "(1) Religious" = h3_models$h3_baseline,
    "(2) Party" = h3_models$h3_party,
    "(3) Military" = h3_models$h3_military,
    "(4) Multi" = h3_models$h3_multi,
    "(5) +Controls" = h3_models$h3_controls,
    "(6) Full" = h3_models$h3_full
  )
)
```

## H4: Support Group Ideology and Democracy Targeting

**H4** tests whether support group ideology shapes target selection among conflict initiators.

```{r}
#| label: tbl-h4
#| tbl-cap: "H4: Support Group Ideology and Democracy Targeting (Initiators Only)"
build_model_table(
  list(
    "(1) Religious" = h4_models$h4_baseline,
    "(2) Party" = h4_models$h4_party,
    "(3) Military" = h4_models$h4_military,
    "(4) Multi" = h4_models$h4_multi,
    "(5) +Controls" = h4_models$h4_controls,
    "(6) Full" = h4_models$h4_full
  )
)
```

## Summary

@tbl-tier1-summary provides a compact summary of the key coefficient estimates and significance levels across H1–H4.

```{r}
#| label: tbl-tier1-summary
#| tbl-cap: "Tier 1 Summary: Key Coefficient Estimates (H1–H4 Full Models)"
build_model_table(
  list(
    "H1: Initiation" = h1_models$h1_full,
    "H2: Targeting" = h2_models$h2_full,
    "H3: Initiation" = h3_models$h3_full,
    "H4: Targeting" = h4_models$h4_full
  )
)
```
